Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-13 doc/manuscript/ESSDmanuscript.Rmd
allometry paragraph

Aboveground biomass, which is typically <200 t C ha^-1^ but can exceed 700 t C ha^-1^ [@anderson-teixeira_carbon_2021], is defined by the IPCC as "all biomass of living vegetation above the soil including stems, stumps, branches, bark, seeds, and foliage" [@ipcc_2006_2006].
IPCC's guidance is that the understory may be excluded the understory if it constitutes a "minor" component, *where quantitative definitions of "understory" and "minor" are not provided*<!--verify-->, but where a commonly applied minimum size sampling threshold for mature forests would be 10 cm stem diameter at breast height (DBH).
A recent study characterizing the contributions of trees in different DBH classes to ecosystem C stocks and fluxes found that trees 1 - 10 cm DBH contributed up to ~8% aboveground biomass, ~17% aboveground woody net primary productivity ($ANPP_{woody.stem}$), and ~20% woody mortality ($M_{woody}$) of mature closed-canopy forests worldwide [@piponiot_distribution_2022].
In regrowth forests, woodlands, or savannas, small trees and shrubs contribute a much larger proportion of C stocks and fluxes [@piponiot_distribution_2022; @refs], and, correspondingly, biomass estimates for these ecosystems tend include smaller size classes [e.g., @refs]. 
Beyond the minimum DBH sampled, forest censuses and biomass estimates also differ in their inclusion of life forms other than dicot trees -- including lianas, ferns, palms, and bamboo -- which in some places can reach large sizes and/or constitute a large fraction of forest C.
*[IPCC guidance]*
Further, it is important to note that this excludes standing dead wood, which is included in remote sensing biomass estimates [@duncanson_aboveground_2021].

@@

A universal challenge in estimating biomass (living or dead) from forest census data is applying appropriate allometries to convert DBH measurements to biomass.
*(Camille/Helene can write this paragraph easily.)*

>>>>

Aboveground biomass, which is typically <200 t C ha^-1^ but can exceed 700 t C ha^-1^ [@anderson-teixeira_carbon_2021], is defined by the IPCC as "all biomass of living vegetation above the soil including stems, stumps, branches, bark, seeds, and foliage" [@ipcc_2006_2006].
IPCC's guidance is that the understory may be excluded the understory if it constitutes a "minor" component, *where quantitative definitions of "understory" and "minor" are not provided*<!--verify-->, but where a commonly applied minimum size sampling threshold for mature forests would be 10 cm stem diameter at breast height (DBH).
A recent study characterizing the contributions of trees in different DBH classes to ecosystem C stocks and fluxes found that trees 1 - 10 cm DBH contributed up to ~8% aboveground biomass, ~17% aboveground woody net primary productivity ($ANPP_{woody.stem}$), and ~20% woody mortality ($M_{woody}$) of mature closed-canopy forests worldwide [@piponiot_distribution_2022].
In regrowth forests, woodlands, or savannas, small trees and shrubs contribute a much larger proportion of C stocks and fluxes [@piponiot_distribution_2022; @refs], and, correspondingly, biomass estimates for these ecosystems tend include smaller size classes [e.g., @refs]. 
Beyond the minimum DBH sampled, forest censuses and biomass estimates also differ in their inclusion of life forms other than dicot trees -- including lianas, ferns, palms, and bamboo -- which in some places can reach large sizes and/or constitute a large fraction of forest C.
*[explain IPCC guidance]*
Further, it is important to note that this excludes standing dead wood, which is included in remote sensing biomass estimates [@duncanson_aboveground_2021].

A universal challenge in estimating biomass (living or dead) from forest census data is applying appropriate allometries to convert DBH measurements to biomass.
Selection of allometries has an enormous influence on estimates of biomass stocks, increments, of fluxes [@clark_landscapescale_2000;@clark_net_2001].
While trusted and standardized allometric equations are becoming increasingly available [@chave_improved_2014;@rejou-mechain_biomass_2017;@gonzalez-akre_allodb_2022], large uncertainties remain.
*[explain IPCC guidance]*

@@

### 2022-04-13 doc/manuscript/ESSDmanuscript.tex
allometry paragraph

Aboveground biomass, which is typically \textless200 t C
ha\textsuperscript{-1} but can exceed 700 t C ha\textsuperscript{-1}
\citep{anderson-teixeira_carbon_2021}, is defined by the IPCC as ``all
biomass of living vegetation above the soil including stems, stumps,
branches, bark, seeds, and foliage'' \citep{ipcc_2006_2006}. IPCC's
guidance is that the understory may be excluded the understory if it
constitutes a ``minor'' component, \emph{where quantitative definitions
of ``understory'' and ``minor'' are not provided}, but where a commonly
applied minimum size sampling threshold for mature forests would be 10
cm stem diameter at breast height (DBH). A recent study characterizing
the contributions of trees in different DBH classes to ecosystem C
stocks and fluxes found that trees 1 - 10 cm DBH contributed up to
\textasciitilde8\% aboveground biomass, \textasciitilde17\% aboveground
woody net primary productivity (\(ANPP_{woody.stem}\)), and
\textasciitilde20\% woody mortality (\(M_{woody}\)) of mature
closed-canopy forests worldwide \citep{piponiot_distribution_2022}. In
regrowth forests, woodlands, or savannas, small trees and shrubs
contribute a much larger proportion of C stocks and fluxes
\citep{piponiot_distribution_2022, refs}, and, correspondingly, biomass
estimates for these ecosystems tend include smaller size classes
\citep[e.g.,][]{refs}. Beyond the minimum DBH sampled, forest censuses
and biomass estimates also differ in their inclusion of life forms other
than dicot trees -- including lianas, ferns, palms, and bamboo -- which
in some places can reach large sizes and/or constitute a large fraction
of forest C. \emph{{[}IPCC guidance{]}} Further, it is important to note
that this excludes standing dead wood, which is included in remote
sensing biomass estimates \citep{duncanson_aboveground_2021}.

@@

A universal challenge in estimating biomass (living or dead) from forest
census data is applying appropriate allometries to convert DBH
measurements to biomass. \emph{(Camille/Helene can write this paragraph
easily.)}

>>>>

Aboveground biomass, which is typically \textless200 t C
ha\textsuperscript{-1} but can exceed 700 t C ha\textsuperscript{-1}
\citep{anderson-teixeira_carbon_2021}, is defined by the IPCC as ``all
biomass of living vegetation above the soil including stems, stumps,
branches, bark, seeds, and foliage'' \citep{ipcc_2006_2006}. IPCC's
guidance is that the understory may be excluded the understory if it
constitutes a ``minor'' component, \emph{where quantitative definitions
of ``understory'' and ``minor'' are not provided}, but where a commonly
applied minimum size sampling threshold for mature forests would be 10
cm stem diameter at breast height (DBH). A recent study characterizing
the contributions of trees in different DBH classes to ecosystem C
stocks and fluxes found that trees 1 - 10 cm DBH contributed up to
\textasciitilde8\% aboveground biomass, \textasciitilde17\% aboveground
woody net primary productivity (\(ANPP_{woody.stem}\)), and
\textasciitilde20\% woody mortality (\(M_{woody}\)) of mature
closed-canopy forests worldwide \citep{piponiot_distribution_2022}. In
regrowth forests, woodlands, or savannas, small trees and shrubs
contribute a much larger proportion of C stocks and fluxes
\citep{piponiot_distribution_2022, refs}, and, correspondingly, biomass
estimates for these ecosystems tend include smaller size classes
\citep[e.g.,][]{refs}. Beyond the minimum DBH sampled, forest censuses
and biomass estimates also differ in their inclusion of life forms other
than dicot trees -- including lianas, ferns, palms, and bamboo -- which
in some places can reach large sizes and/or constitute a large fraction
of forest C. \emph{{[}explain IPCC guidance{]}} Further, it is important
to note that this excludes standing dead wood, which is included in
remote sensing biomass estimates \citep{duncanson_aboveground_2021}.

A universal challenge in estimating biomass (living or dead) from forest
census data is applying appropriate allometries to convert DBH
measurements to biomass. Selection of allometries has an enormous
influence on estimates of biomass stocks, increments, of fluxes
\citep{clark_landscapescale_2000, clark_net_2001}. While trusted and
standardized allometric equations are becoming increasingly available
\citep{chave_improved_2014, rejou-mechain_biomass_2017, gonzalez-akre_allodb_2022},
large uncertainties remain. \emph{{[}explain IPCC guidance{]}}

@@
